---
title: "Principal Component Analysis: get the top 100 PCs"
date: "Created: 2021-01-23; updated: `r Sys.Date()`"
author: 
  - name: "Hua Zou"
    email: "zouhua1@outlook.com"
output: 
  html_notebook:
    codes: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(ggplot2)
library(data.table)
library(factoextra)

rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)
```

## What is PCA

Principal component analysis is an unsupervised method to find the greatest sources of variation from the multiple dimensions data. It used the partial least squares to perform the covariance decomposition.


### load data
```{r}
phen <- fread("../../Result/phenotype/common_survival_data.tsv")
all_prf <- fread("../../Result/profile/All_filter_merge_profile.tsv")
```

### get the principal components by PCA

* Row: samples

* Column: omics-data features

Before performing PCA, we need scale the dataset, and use the built-in prcomp function implemented singular value decomposition algorithm.

```{r}
get_pc_scores <- function(dataset = all_prf){
  
  # dataset = all_prf
  
  sid <- intersect(phen$Barcode, dataset$SampleID) 
  prof <- dataset %>% filter(SampleID%in%sid) %>%
    column_to_rownames("SampleID")
  
  fit <- prcomp(prof, scale. = T, center = T)
  
  # rotation <- fit$rotation
  scores <- fit$x
  
  return(scores)
}
```


### Run
```{r}
pc_scores <- get_pc_scores(dataset = all_prf)

pc_scores_top100 <- pc_scores %>% data.frame() %>% 
  dplyr::select(c(1:100)) %>% 
  rownames_to_column("SampleID")

write.csv(data.frame(pc_scores) %>% rownames_to_column("SampleID"),
          "../../Result/profile/PCA_all_features.csv", row.names = F)

write.csv(pc_scores_top100,
          "../../Result/profile/PCA_top100_features.csv", row.names = F)
```


### version
```{r}
sessionInfo()
```


### Reference 

1. [Articles - Principal Component Methods in R: Practical Guide](http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/)

2. [Principal Component Analysis in R](https://www.r-bloggers.com/2017/01/principal-component-analysis-in-r/)



