---
title: "Enriched pathway analysis: based on cluterProfiler and enrichplot"
date: "Created: 2021-01-28 Updated: `r Sys.Date()`"
author: 
  - name: "Hua Zou"
    email: "zouhua1@outlook.com"
output: 
  html_notebook:
    codes: hide
---

## Introduction

Performing Enriched pathway analysis via **clusterProfiler** and **enrichplot** package based on the DESeq2 results

* *clusterProfiler* calculates the enrichment score

* *enrichplot* displays the enrichment result

Two approaches for enrichment analysis:

* Over Representation Analysis: the up- or down-regulated DEGs 

* Gene Set Enrichment Analysis: All the ranked DEGs

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(ggplot2)
library(data.table)
library(clusterProfiler)
# library(enrichplot)
library(patchwork)

rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)

grp <- c("S1", "S2")
grp.col <- c("#6C326C", "#77A2D1")
```


### load data 

* gene Expression DESeq2 object
```{r}
phen <- fread("../../Result/phenotype/phenotype_cluster.csv")
geneExp <- readRDS("../../Result/Differential/geneExp_Diff.RDS")
```


### Curation Function 

* Over Representation Analysis: the DEgenes vectors (up or down)

* Gene Set Enrichment Analysis: all the ranked genes by statistics
```{r}
# ORA 
get_ORA <- function(dataset=geneExp$diff_res,
                    group=grp[1],
                    type="kegg"){
  
  # dataset=geneExp$diff_res
  # group=grp[1]
  # type="kegg"
  
  # extract DEGs
  dat <- dataset %>% data.frame() %>%
    rownames_to_column("SYMBOL") %>%
    filter(Enrichment%in%group) %>%
    distinct() %>%
    arrange(log2FoldChange)
  
  # transform geneid
  require(org.Hs.eg.db)
  symbol2entrz <- AnnotationDbi::select(org.Hs.eg.db,
                                      key=dat$SYMBOL, 
                                      columns="ENTREZID",
                                      keytype="SYMBOL")
  symbol2entrz <- as_tibble(symbol2entrz)

  dat_v2 <- inner_join(symbol2entrz, dat, by ="SYMBOL") %>%
    dplyr::select(ENTREZID, log2FoldChange) %>%
    na.omit() %>% 
    distinct() %>% 
    group_by(ENTREZID) %>% 
    summarize(log2FoldChange=mean(log2FoldChange)) %>%
    ungroup() %>%
    arrange(desc(log2FoldChange))
  
  geneList <- tibble::deframe(dat_v2)  
  if(type == "kegg"){
    es_res <- enrichKEGG(gene          = names(geneList),
                         organism      = "hsa",
                         pAdjustMethod = "BH",
                         pvalueCutoff  = 0.01,
                         qvalueCutoff  = 0.05)
  }else if(type == "GO"){
    es_res <- enrichGO(gene          = names(geneList),
                       OrgDb         = org.Hs.eg.db,
                       keyType       = "ENTREZID",
                       ont           = "CC",
                       pAdjustMethod = "BH",
                       pvalueCutoff  = 0.01,
                       qvalueCutoff  = 0.05)    
  }
  
  res <- list(glist=geneList, es=es_res) 
  
  return(res)
}

# GSEA
get_GSEA <- function(dataset=geneExp$diff_res,
                     type="kegg"){
  
  # dataset=geneExp$diff_res
  # type="kegg"
  
  dat <- dataset %>% data.frame() %>%
    rownames_to_column("SYMBOL") 
  
  # transform geneid
  require(org.Hs.eg.db)
  symbol2entrz <- AnnotationDbi::select(org.Hs.eg.db,
                                        key=dat$SYMBOL, 
                                        columns="ENTREZID",
                                        keytype="SYMBOL")
  symbol2entrz <- as_tibble(symbol2entrz)

  dat_v2 <- inner_join(symbol2entrz, dat, by ="SYMBOL") %>%
    dplyr::select(ENTREZID, log2FoldChange) %>%
    na.omit() %>% 
    distinct() %>% 
    group_by(ENTREZID) %>% 
    summarize(log2FoldChange=mean(log2FoldChange)) %>%
    ungroup() %>%
    arrange(desc(log2FoldChange))
  
  geneList <- tibble::deframe(dat_v2)
  require(fgsea)
  if(type == "kegg"){
    es_res <- gseKEGG(gene          = geneList,
                      organism      = "hsa",
                      keyType       = "kegg",
                      nPerm         = 1000,
                      minGSSize     = 10,
                      maxGSSize     = 500,
                      pAdjustMethod = "BH",
                      pvalueCutoff  = 0.05,
                      by            = "fgsea")
  }else if(type == "GO"){
    es_res <- gseGO(gene          = geneList,
                    OrgDb         = org.Hs.eg.db,
                    ont           = "CC",
                    nPerm         = 1000,
                    minGSSize     = 10,
                    maxGSSize     = 500,                 
                    pAdjustMethod = "BH",
                    pvalueCutoff  = 0.05)    
  }
  
  res <- list(glist=geneList, es=es_res) 
  
  return(res)
}

plot_Escore <- function(dataset=GSEA_kegg,
                        type="barplot",
                        number=10,
                        pathway="Oxidative phosphorylation",
                        pathid="hsa04110"){
  # dataset=GSEA_kegg
  # type="barplot"
  # number=10
  # pathway="Oxidative phosphorylation"
  # pathid="hsa04110"
  
  my_theme <- theme(axis.title = element_text(face = 'bold',color = 'black',size = 14),
                  axis.text = element_text(color = 'black',size = 10),
                  text = element_text(size = 8, color = "black", family="serif"),
                  legend.position = 'right',
                  legend.key.height = unit(0.6,'cm'),
                  legend.text = element_text(face = "bold", color = 'black',size = 10),
                  strip.text = element_text(face = "bold", size = 14))
  
  edox <- setReadable(dataset$es, 'org.Hs.eg.db', 'ENTREZID')
  
  if(type == "barplot"){
    pl <- barplot(edox, showCategory = number)+
      my_theme
  }else if(type == "dotplot"){
    pl <- enrichplot::dotplot(edox, showCategory=number)+
      my_theme
  }else if(type == "network"){
    pl <- enrichplot::cnetplot(edox, categorySize="pvalue", foldChange=dataset$glist, showCategory = 3)+
      my_theme
  }else if(type == "heatmap"){
    pl <- enrichplot::heatplot(edox, foldChange=dataset$glist, showCategory = number)+
      my_theme
  }else if(type == "enrichment"){
    pl <- enrichplot::emapplot(dataset$es, pie_scale=1.5, layout="kk", showCategory = number)+
      my_theme
  }else if(type == "UpSet"){
    require(ggupset)
    pl <- enrichplot::upsetplot(dataset$es, n=number)+
      my_theme
  }else if(type == "GSEA"){
    npath <- pmatch(pathway, dataset$es$Description)
    pl <- enrichplot::gseaplot2(dataset$es, geneSetID = npath, title = dataset$es$Description[npath])+
      my_theme
  }else if(type == "pubmed"){
    terms <- dataset$es$Description[1:number]
    pl <- enrichplot::pmcplot(terms, 2010:2021, proportion = T)+
      my_theme
  }else if(type == "pathview"){
    require(pathview)
    pl <- pathview(gene.data  = dataset$glist,
                   pathway.id = pathid,
                   species    = "hsa",
                   limit      = list(gene=max(abs(dataset$glist)), cpd=1))    
  }
  return(pl)
}
```


### Over Representation Analysis

* Calculate the Enrichment results
```{r}
# KEGG 
DEGs_ORA_kegg <- get_ORA(dataset = geneExp$diff_res, group = grp, type = "kegg")
DEGs_ORA_kegg_S1 <- get_ORA(dataset = geneExp$diff_res, group = grp[1], type = "kegg")
DEGs_ORA_kegg_S2 <- get_ORA(dataset = geneExp$diff_res, group = grp[2], type = "kegg")

# GO
DEGs_ORA_GO <- get_ORA(dataset = geneExp$diff_res, group = grp, type = "GO")
DEGs_ORA_GO_S1 <- get_ORA(dataset = geneExp$diff_res, group = grp[1], type = "GO")
DEGs_ORA_GO_S2 <- get_ORA(dataset = geneExp$diff_res, group = grp[2], type = "GO")

save(DEGs_ORA_kegg, DEGs_ORA_kegg_S1, DEGs_ORA_kegg_S2, 
     DEGs_ORA_GO, DEGs_ORA_GO_S1, DEGs_ORA_GO_S2,
     file = "../../Result/Enrichment/DEGs_ORA_KEGG_GO.RData")
```

* Visualization : KEGG barplot 
```{r, fig.width=16, fig.height=8}
DEGs_ORA_kegg_barplot <- plot_Escore(dataset = DEGs_ORA_kegg, number = 10, type = "barplot")
DEGs_ORA_kegg_barplot_S1 <- plot_Escore(dataset = DEGs_ORA_kegg_S1, number = 10, type = "barplot")
DEGs_ORA_kegg_barplot_S2 <- plot_Escore(dataset = DEGs_ORA_kegg_S2, number = 10, type = "barplot")

(DEGs_ORA_kegg_barplot | DEGs_ORA_kegg_barplot_S1 | DEGs_ORA_kegg_barplot_S2)+
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = "A")
```

* Visualization : KEGG dotplot
```{r, fig.width=16, fig.height=8}
DEGs_ORA_kegg_dotplot <- plot_Escore(dataset = DEGs_ORA_kegg, number = 10, type = "dotplot")
DEGs_ORA_kegg_dotplot_S1 <- plot_Escore(dataset = DEGs_ORA_kegg_S1, number = 10, type = "dotplot")
DEGs_ORA_kegg_dotplot_S2 <- plot_Escore(dataset = DEGs_ORA_kegg_S2, number = 10, type = "dotplot")

(DEGs_ORA_kegg_dotplot | DEGs_ORA_kegg_dotplot_S1 | DEGs_ORA_kegg_dotplot_S2)+
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = "A")
```

* Visualization : KEGG network
```{r, fig.width=16, fig.height=12}
DEGs_ORA_kegg_network <- plot_Escore(dataset = DEGs_ORA_kegg, number = 4, type = "network")
DEGs_ORA_kegg_network_S1 <- plot_Escore(dataset = DEGs_ORA_kegg_S1, number = 4, type = "network")
DEGs_ORA_kegg_network_S2 <- plot_Escore(dataset = DEGs_ORA_kegg_S2, number = 4, type = "network")

(DEGs_ORA_kegg_network | DEGs_ORA_kegg_network_S1 | DEGs_ORA_kegg_network_S2)+
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = "A")
```

* Visualization : KEGG Enrichment Map
```{r, fig.width=16, fig.height=12}
DEGs_ORA_kegg_enrichment <- plot_Escore(dataset = DEGs_ORA_kegg, number = 10, type = "enrichment")
DEGs_ORA_kegg_enrichment_S1 <- plot_Escore(dataset = DEGs_ORA_kegg_S1, number = 10, type = "enrichment")
DEGs_ORA_kegg_enrichment_S2 <- plot_Escore(dataset = DEGs_ORA_kegg_S2, number = 10, type = "enrichment")

(DEGs_ORA_kegg_enrichment | DEGs_ORA_kegg_enrichment_S1 | DEGs_ORA_kegg_enrichment_S2)+
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = "A")
```


### Gene Set Enrichment Analysis

* calculate enrichment score
```{r, fig.width=15}
GSEA_kegg <- get_GSEA(dataset = geneExp$diff_res, type = "kegg")
GSEA_GO <- get_GSEA(dataset = geneExp$diff_res, type = "GO")
save(GSEA_kegg, GSEA_GO,
     file = "../../Result/Enrichment/GSEA_KEGG_GO.RData")
```

* visualization: KEGG dotplot
```{r}
GSEA_kegg_dotplot <- plot_Escore(dataset = GSEA_kegg, type = "dotplot", number = 20)
GSEA_kegg_dotplot

if(!dir.exists("../../Result/figure/KEGG")){
  dir.create("../../Result/figure/KEGG", recursive = T)
}
ggsave("../../Result/figure/KEGG/GSEA_KEGG_barplot.pdf", GSEA_kegg_dotplot, width = 8, height = 6)
```

* visualization: KEGG Gene-Concept Network
```{r}
GSEA_kegg_network <- plot_Escore(dataset = GSEA_kegg, type = "network", number = 20)
GSEA_kegg_network
ggsave("../../Result/figure/KEGG/GSEA_KEGG_network.pdf", GSEA_kegg_network, width = 8, height = 8)
```

* visualization: KEGG Enrichment Map
```{r}
GSEA_kegg_enrichment <- plot_Escore(dataset = GSEA_kegg, type = "enrichment", number = 20)
GSEA_kegg_enrichment
ggsave("../../Result/figure/KEGG/GSEA_KEGG_enrichment.pdf", GSEA_kegg_enrichment, width = 8, height = 8)
```

* visualization: KEGG running score and preranked list of GSEA result
```{r}
# Oxidative phosphorylation
GSEA_kegg_Oxidative_phosphorylation <- plot_Escore(dataset = GSEA_kegg, type = "GSEA", pathway = "Oxidative phosphorylation")
GSEA_kegg_Oxidative_phosphorylation
ggsave("../../Result/figure/KEGG/GSEA_KEGG_Oxidative_phosphorylation.pdf", GSEA_kegg_Oxidative_phosphorylation, width = 8, height = 8)

# Ribosome
GSEA_kegg_Ribosome <- plot_Escore(dataset = GSEA_kegg, type = "GSEA", pathway = "Ribosome")
GSEA_kegg_Ribosome
ggsave("../../Result/figure/KEGG/GSEA_KEGG_Ribosome.pdf", GSEA_kegg_Ribosome, width = 8, height = 8)

# Cytokine-cytokine receptor interaction
GSEA_kegg_Cytokine <- plot_Escore(dataset = GSEA_kegg, type = "GSEA", pathway = "Cytokine-cytokine receptor interaction")
GSEA_kegg_Cytokine
ggsave("../../Result/figure/KEGG/GSEA_KEGG_Cytokine.pdf", GSEA_kegg_Cytokine, width = 8, height = 8)
```

* visualization: KEGG pathway view
```{r}
# Cytokine-cytokine receptor interaction
plot_Escore(dataset = GSEA_kegg, type = "pathview", pathid = "hsa04060")
```

### version
```{r}
sessionInfo()
```


### Reference 

1. [Functional Enrichment Analysis Methods](http://yulab-smu.top/clusterProfiler-book/chapter2.html) 

2. [Visualization of Functional Enrichment Result](http://yulab-smu.top/clusterProfiler-book/chapter12.html#enrichment-map)
