---
title: "survival prediction on clinical features"
date: "Created: 2021-01-30; updated: `r Sys.Date()`"
author: 
  - name: "Hua Zou"
    email: "zouhua1@outlook.com"
output: 
  html_notebook:
    codes: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(stringr)
library(survival)
library(data.table)
library(glmnet)
library(c060)
library(peperr)
library(Hmisc)
library(plotrix)


rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)
```


## Introduciton

> The Cox proportional hazards model is commonly used for the study of the relationship beteween predictor variables and survival time and penalize the negative log of the partial likelihood with an elastic net penalty via glmnet R package.

### load data
```{r}
phen <- fread("../../Result/phenotype/phenotype_cluster.csv")
```


### Function

* sole clinical features

* clinical features + autoencoder labels
```{r}
get_clinical <- function(x=phen, 
                         label=TRUE){
  
  # x=phen
  # label=TRUE
  
  # pathologic_T 
  x$Te1 <- str_extract(x$TNM, "T\\d[a-z]+")
  x$Te2 <- str_extract(x$TNM, "T\\d")
  x$Te1 <- with(x, ifelse(is.na(Te1), Te2, Te1))
  x$Te2[x$Te2 == "T1"] <- 1
  x$Te2[x$Te2 == "T2"] <- 2
  x$Te2[x$Te2 == "T3"] <- 3
  x$Te2[x$Te2 == "T4"] <- 4
  
  # pathologic_N
  x$N <- str_extract(x$TNM, "N\\d")
  x$N <- str_extract(x$N, "\\d")
  x$N[is.na(x$N)] <- 2
  
  # pathologic_M
  x$M <- str_extract(x$TNM, "M\\d")
  x$M <- str_extract(x$M, "\\d")
  x$M[is.na(x$M)] <- 2
  
  # Gender
  x$Gender <- ifelse(x$Gender == "MALE", 1, 2)
  
  # stage
  x$Stage1 <- str_trim(str_extract(x$Stage, "\\s[H-Z]+"),
                              side = c("both", "left", "right"))
  x$Stage1[x$Stage1 == "I"] <- 1
  x$Stage1[x$Stage1 == "II"] <- 2
  x$Stage1[x$Stage1 == "III"] <- 3
  x$Stage1[x$Stage1 == "IV"] <- 4
  x$Stage1[is.na(x$Stage1)] <- 5
  
  x$Race <- as.numeric(factor(x$Race))
  x$cancer_status <- as.numeric(factor(x$cancer_status))
  x$laterality <- as.numeric(factor(x$laterality))
  x$Grade <- as.numeric(factor(x$Grade))
  x$Te2 <- as.numeric(factor(x$Te2))
  x$N <- as.numeric(factor(x$N))
  x$M <- as.numeric(factor(x$M))
  x$Stage1 <- as.numeric(factor(x$Stage1))
  x$Cluster <- as.numeric(factor(x$Cluster)) 
  
  temp <- x %>% dplyr::select(-c("Stage", "TNM","Te1")) %>%
    dplyr::rename(Stage=Stage1, T=Te2, time=OS.Time, status=OS) %>%
    filter(time != 0) %>%
    column_to_rownames("Barcode")
  
  if(label){
    res <- temp
  }else{
    res <- temp %>% dplyr::select(-Cluster)
  }
  return(res)  
}

# get predict matrix
get_matrix <- function(dataset=clinical_params, 
                       nonzero.coef=NULL){

  # dataset=clinical_params
  # nonzero.coef=NULL

  surv_obj <- Surv(dataset$time, dataset$status) 
  predict_matrix <- model.matrix(~ ., data=dataset, 
              contrasts.arg=lapply(dataset[, sapply(dataset, is.factor)],contrasts))
  predict_matrix <- subset(predict_matrix, select=c(-time, -status))
 
  if(!is.null(nonzero.coef)){
    predict_matrix <- predict_matrix[, nonzero.coef]
  }
 
  res <- list(predict_matrix=predict_matrix,
              surv_obj=surv_obj)
  return(res)
}

# glmnet for cox-ph model
glmnetfun <- function(predict_matrix=matrix_res$predict_matrix,
                      surv_obj=matrix_res$surv_obj,
                      alpha=0){
  
  # predict_matrix=matrix_res$predict_matrix
  # surv_obj=matrix_res$surv_obj
  # alpha=0
  
  cv.fit <- cv.glmnet(x=predict_matrix, 
                      y=surv_obj, 
                      family="cox",
                      alpha = alpha,
                      type.measure = "C")
 
  glmnet.obj <- cv.fit$glmnet.fit
  optimal.lambda <- cv.fit$lambda.min  
  lambda.index <- which(cv.fit$lambda == optimal.lambda) 
  optimal.beta <- glmnet.obj$beta[, lambda.index] 
  
  # take only covariates for which beta is not zero   
  nonzero.coef <- abs(optimal.beta) > 0 
  selectedBeta <- optimal.beta[nonzero.coef] 
  selectedVar <- predict_matrix[, nonzero.coef] 
  reformat_dataSet <- as.data.frame(cbind(surv_obj, selectedVar))
 
  # create coxph object with pre-defined coefficients 
  glmnet.cox <- coxph(Surv(time, status) ~ ., data=reformat_dataSet, init=selectedBeta, iter=200)

  res <- list(glmnet.cv=cv.fit, 
              glmnet.obj=glmnet.obj, 
              glmnet.cox=glmnet.cox, 
              nonzero.coef=nonzero.coef, 
              selectedVar=selectedVar, 
              reformat_dataSet=reformat_dataSet)
  return(res)
}

# Calculates Brier score
get_Brier_score <- function(predict_matrix=matrix_res$predict_matrix, 
                            surv_obj=matrix_res$surv_obj, 
                            model=c("cox","glmnet"),
                            nfolds=10){
  
  # predict_matrix=matrix_res$predict_matrix
  # surv_obj=matrix_res$surv_obj
  # model="glmnet"
  # nfolds=10
  
  if(model == "cox"){
    peperr <- peperr.cox <- peperr(response=surv_obj, 
                                   x=predict_matrix,
                                   fit.fun=fit.coxph, 
                                   load.all=TRUE,
                                   indices=resample.indices(n=nrow(surv_obj), 
                                                            method="boot",
                                                            sample.n=50))      
  }else if(model == "glmnet"){
    peperr <- peperr(response=surv_obj, 
                     x=predict_matrix,
                     fit.fun=fit.glmnet, 
                     args.fit=list(family="cox"),
                     complexity=complexity.glmnet,
                     args.complexity=list(family="cox"), 
                     load.all=TRUE,
                     indices=resample.indices(n=nrow(surv_obj), 
                                              method="boot", 
                                              sample.n=nfolds))
  }
 
  # Get error predictions from peperr
  prederr <- perr(peperr)
 
  # Integrated prediction error Brier score calculation
  ipec <- ipec(prederr, eval.times=peperr$attribute, response=surv_obj)
 
  res <- list(prederr=prederr, ipec=ipec)
  return(res)
}

# Calculate c-index (Harrell's)
get_Cindex <- function(dataset=clinical_params, 
                       nonzero.coef=NULL,
                       iteration=10){

  # dataset=clinical_params
  # nonzero.coef=NULL
  # iteration=10

  
  i <- 1
  train_results_list <- list()
  cIndexCox <- list()
  cIndexglmnet <- list()   
   
  while (i < iteration){
    
    print(paste("Iteration:",i))
    train <- sample(1:nrow(dataset), nrow(dataset), replace = TRUE)
    trainset <- dataset[train, ]
    testset <- dataset[-train, ]
    train_matrix <- get_matrix(trainset)
   
    train_results <- glmnetfun(train_matrix$predict_matrix, train_matrix$surv_obj) 
    train_results_list <- c(train_results_list, list(train_results))
   
    # compute c-index (Harrell's) for cox-glmnet models
    if (!is.null(train_results$glmnet.cox)){
      # creates a matrix of only the variables identified by the glmnet
      test_matrix <- get_matrix(testset, nonzero.coef=train_results$nonzero.coef)
      reformat_testSet <- as.data.frame(cbind(test_matrix$surv_obj, test_matrix$predict_matrix))
      rcorr_cens <- rcorr.cens(predict(train_results$glmnet.cox, 
                                       reformat_testSet), 
                               test_matrix$surv_obj)
      
      cIndexCox <- c(cIndexCox, 1 - rcorr_cens[1])
      i <- i + 1
      # compute c-index (Harrell's) for glmnet models
      test_matrix <- get_matrix(testset)
      rcorr_cens <- rcorr.cens(predict(train_results$glmnet.cv,                                             
                                       test_matrix$predict_matrix),
                               test_matrix$surv_obj)
      cIndexglmnet <- c(cIndexglmnet, 1 - rcorr_cens[1])
    }
  }  
  
  meanCICox <- round(c(mean(unlist(cIndexCox), na.rm=TRUE), 
                 plotrix::std.error(unlist(cIndexCox))), 2)
  meanCIglmnet <- round(c(mean(unlist(cIndexglmnet), na.rm=TRUE), 
                    plotrix::std.error(unlist(cIndexglmnet))), 2)
  
  res <- list(Cox=paste(meanCICox, collapse = "+/-"),
              glmnet=paste(meanCIglmnet, collapse = "+/-"))
  return(res)
}
```

### prediction

* with Cluster labels
```{r}
clinical_params <- get_clinical(x = phen)
matrix_res <- get_matrix(dataset = clinical_params)

Bscore <- get_Brier_score(predict_matrix = matrix_res$predict_matrix,
                surv_obj = matrix_res$surv_obj)
mean(Bscore$prederr)

get_Cindex(dataset = clinical_params, iteration = 10)
```

* without Cluster labels
```{r}
clinical_params_no <- get_clinical(x = phen, label = FALSE)
matrix_res_no <- get_matrix(dataset = clinical_params_no)

Bscore_no <- get_Brier_score(predict_matrix = matrix_res_no$predict_matrix,
                surv_obj = matrix_res_no$surv_obj)
mean(Bscore_no$prederr)

get_Cindex(dataset = clinical_params_no, iteration = 10)
```


### version
```{r}
sessionInfo()
```


### Reference 

1. [Regularized Cox Regression](https://glmnet.stanford.edu/articles/Coxnet.html) 

2. [How and when: ridge regression with glmnet](https://drsimonj.svbtle.com/ridge-regression-with-glmnet)

3. [一致性指数：Harrell’concordance index:C-index](https://www.jianshu.com/p/5e648f0f49ed)

4. [Calculate Concordance index](https://glmnet.stanford.edu/reference/Cindex.html)

5. [Calculate Brier score on glmnet object](https://stats.stackexchange.com/questions/63530/problem-with-bootcv-for-coxph-in-pec-after-feature-selection-with-glmnet-lasso)

6. [Calculate Brier score on glmnet object](https://pastebin.com/yNpZMzNc)
