---
title: "survival prediction on clinical features"
date: "Created: 2021-01-23; updated: `r Sys.Date()`"
author: 
  - name: "Hua Zou"
    email: "zouhua1@outlook.com"
output: 
  html_notebook:
    codes: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(stringr)
library(survival)
library(data.table)
library(glmnet)
library(c060)
library(peperr)
library(Hmisc)

rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)
```


## Introduciton

> The Cox proportional hazards model is commonly used for the study of the relationship beteween predictor variables and survival time and penalize the negative log of the partial likelihood with an elastic net penalty via glmnet R package.

### load data
```{r}
phen <- fread("../../Result/phenotype/phenotype_cluster.csv")
```


### Function

* sole clinical features

* clinical features + autoencoder labels
```{r}
get_clinical <- function(x=phen){
  
  # x=phen
  
  # pathologic_T 
  x$Te1 <- str_extract(x$TNM, "T\\d[a-z]+")
  x$Te2 <- str_extract(x$TNM, "T\\d")
  x$Te1 <- with(x, ifelse(is.na(Te1), Te2, Te1))
  x$Te2[x$Te2 == "T1"] <- 1
  x$Te2[x$Te2 == "T2"] <- 2
  x$Te2[x$Te2 == "T3"] <- 3
  x$Te2[x$Te2 == "T4"] <- 4
  
  # pathologic_N
  x$N <- str_extract(x$TNM, "N\\d")
  x$N <- str_extract(x$N, "\\d")
  x$N[is.na(x$N)] <- 2
  
  # pathologic_M
  x$M <- str_extract(x$TNM, "M\\d")
  x$M <- str_extract(x$M, "\\d")
  x$M[is.na(x$M)] <- 2
  
  # Gender
  x$Gender <- ifelse(x$Gender == "MALE", 1, 2)
  
  # stage
  x$Stage1 <- str_trim(str_extract(x$Stage, "\\s[H-Z]+"),
                              side = c("both", "left", "right"))
  x$Stage1[x$Stage1 == "I"] <- 1
  x$Stage1[x$Stage1 == "II"] <- 2
  x$Stage1[x$Stage1 == "III"] <- 3
  x$Stage1[x$Stage1 == "IV"] <- 4
  x$Stage1[is.na(x$Stage1)] <- 5
  
  # remove NA value 
  # x$N <- ifelse(is.na(x$N), "Unknown", x$N)
  # x$M <- ifelse(is.na(x$M), "Unknown", x$M)
  # x$Stage1 <- ifelse(is.na(x$Stage1), "Unknown", x$Stage1)
  # x$Grade <- ifelse(x$Grade%in%c("[Not Available]", "GX"), "Unknown", x$Grade)  
  # x$cancer_status <- ifelse(x$cancer_status%in%c("[Not Available]", "[Discrepancy]"), 
  #                           "Unknown", x$cancer_status)
  # x$Race <- ifelse(x$Race == "[Not Available]", "Unknown", x$Race)
  # tmp <- x %>% filter(N != "Unknown") %>%
  #   filter(M != "Unknown") %>%
  #   filter(Stage1 != "Unknown") %>%
  #   filter(Grade != "Unknown") %>%
  #   filter(cancer_status != "Unknown") %>%
  #   filter(Race != "Unknown")
  
  x$Race <- as.numeric(factor(x$Race))
  x$cancer_status <- as.numeric(factor(x$cancer_status))
  x$laterality <- as.numeric(factor(x$laterality))
  x$Grade <- as.numeric(factor(x$Grade))
  x$Te2 <- as.numeric(factor(x$Te2))
  x$N <- as.numeric(factor(x$N))
  x$M <- as.numeric(factor(x$M))
  x$Stage1 <- as.numeric(factor(x$Stage1))
  x$Cluster <- as.numeric(factor(x$Cluster)) 
  
  res <- x %>% dplyr::select(-c("Stage", "TNM","Te1")) %>%
    dplyr::rename(Stage=Stage1, T=Te2, time=OS.Time, status=OS) 
  
  return(res)  
}

get_index_score <- function(x=clinical_params,
                            label=TRUE){
  
  # x=get_clinical(x = phen)
  # label=TRUE
  
  size <- round(nrow(x)*0.8)
  sid <- sample(x$Barcode, size)
  
  trainData <- x %>% filter(Barcode%in%sid) %>% column_to_rownames("Barcode")
  testData <- x %>% filter(!Barcode%in%sid) %>% column_to_rownames("Barcode")
  
  if(label){
    train_phen <- trainData %>% dplyr::select(time, status) %>% as.matrix()
    train_prof <- subset (trainData, select=c(-time,-status)) %>% as.matrix()
    
    test_phen <- testData %>% dplyr::select(time, status) %>% as.matrix()
    test_prof <- subset (testData, select=c(-time,-status)) %>% as.matrix()    
  }else{
    trainData_v2 <- trainData %>% dplyr::select(-Cluster)
    testData_v2 <- testData %>% dplyr::select(-Cluster)   
    train_phen <- trainData_v2 %>% dplyr::select(time, status) %>% as.matrix()
    train_prof <- subset (trainData_v2, select=c(-time,-status)) %>% as.matrix()
    
    test_phen <- testData_v2 %>% dplyr::select(time, status) %>% as.matrix()
    test_prof <- subset (testData_v2, select=c(-time,-status)) %>% as.matrix()     
  }
  
  set.seed(123)
  cv.fit <- cv.glmnet(x=train_prof, 
                      y=train_phen, 
                      family = "cox", 
                      type.measure = "C", # Harrel's concordance
                      nfolds = 10,
                      alpha = 0,
                      nlambda = 100)
  # plot(cv.fit)  
  
  # build model
  fit <- glmnet(x=train_prof, 
                y=train_phen, 
                family = "cox", 
                alpha = 0,
                lambda = cv.fit$lambda.min) # For a more parsimoneous model use lambda.1se
  # coef(fit)
  
  # C-index
  pred <- predict(fit, newx = test_prof)
  cindex <- apply(pred, 2, Cindex, y=test_phen)
  
  res <- list(cv_fit=cv.fit, fit=fit, cindex=cindex)
  return(res)
}
```

### prediction

* with Cluster labels
```{r}
clinical_params <- get_clinical(x = phen)

sur_label <- get_index_score(x = clinical_params, label = TRUE)
# cv fit 
plot(sur_label$cv_fit)
# fit
coef(sur_label$fit)
# Cindex 
sur_label$cindex
```

* without Cluster labels
```{r}
sur_nolabel <- get_index_score(x = clinical_params, label = FALSE)
# cv fit 
plot(sur_nolabel$cv_fit)
# fit
coef(sur_nolabel$fit)
# Cindex 
sur_nolabel$cindex
```


### version
```{r}
sessionInfo()
```


### Reference 

1. [Regularized Cox Regression](https://glmnet.stanford.edu/articles/Coxnet.html) 

2. [How and when: ridge regression with glmnet](https://drsimonj.svbtle.com/ridge-regression-with-glmnet)

3. [一致性指数：Harrell’concordance index:C-index](https://www.jianshu.com/p/5e648f0f49ed)

4. [Calculate Concordance index](https://glmnet.stanford.edu/reference/Cindex.html)

5. [Calculate Brier score on glmnet object](https://stats.stackexchange.com/questions/63530/problem-with-bootcv-for-coxph-in-pec-after-feature-selection-with-glmnet-lasso)

6. [Calculate Brier score on glmnet object](https://pastebin.com/yNpZMzNc)

### the codes from the network
```{r}
library(survival)
library(glmnet)
library(c060)
library(peperr)
library(Hmisc)
##-----------------------------------------------------------------------------------
##                                    Accessory Functions
##-----------------------------------------------------------------------------------
## 
## This function takes a predictor matrix and a Surv obj and uses the glmnet lasso 
## regularization method to select the most predictive features and create a coxph 
## object predict_matrix is the 'dataset' reformated to replace categorical variables
## to binary with dummy
 
dataset2predictmatrix <- function(dataset, 
                                  nonzero.coef=NULL){
  
  # creat Y (survival matrix) for glmnet
  surv_obj <- Surv(dataset$time,dataset$status) 
 
  # tranform categorical variables into binary variables with dummy for dataset
  predict_matrix <- model.matrix(~ ., data=dataset, 
                     contrasts.arg=lapply(dataset[,sapply(dataset, is.factor)], 
                                          contrasts))
 
  # remove the statu/time variables from the predictor matrix (x) for glmnet
  predict_matrix <- subset(predict_matrix, select=c(-time,-status))
 
  if(!is.null(nonzero.coef)){
    predict_matrix <- predict_matrix[, nonzero.coef]
  }
 
  res <- list(predict_matrix=predict_matrix, surv_obj=surv_obj)
  return(res)
}
 
glmnetfun <- function(predict_matrix, surv_obj){
 
  # create a glmnet cox object using lasso regularization and cross validation
  glmnet.cv <- cv.glmnet(predict_matrix, surv_obj, family="cox")
 
  # get the glmnet model 
  glmnet.obj <- glmnet.cv$glmnet.fit
 
  # find lambda index for the models with least partial likelihood deviance (by cv.glmnet) 
  optimal.lambda <- glmnet.cv$lambda.min    # For a more parsimoneous model use lambda.1se     
  lambda.index <- which(glmnet.obj$lambda == optimal.lambda) 
 
  # take beta for optimal lambda 
  optimal.beta <- glmnet.obj$beta[, lambda.index] 
 
  # find non zero beta coef 
  nonzero.coef <- abs(optimal.beta) > 0 
  selectedBeta <- optimal.beta[nonzero.coef] 
 
  # take only covariates for which beta is not zero 
  selectedVar <- predict_matrix[, nonzero.coef] 
 
  # create a dataframe for trainSet with time, status and selected variables in binary representation for evaluation in pec
  reformat_dataSet <- as.data.frame(cbind(surv_obj, selectedVar))
 
  # create coxph object with pre-defined coefficients 
  glmnet.cox <- coxph(Surv(time,status) ~ ., data=reformat_dataSet, init=selectedBeta, iter=200)
 
  # # glmnet.cox only with meaningful features selected by stepwise bidirectional AIC feature selection
  # glmnet.cox.meaningful <- step(coxph(Surv(time,status) ~ ., data=reformat_dataSet), direction="both")
  # selectedVarCox <- selectedVar[, attr(glmnet.cox.meaningful$terms, "term.labels")]
  
  # Returned object -> list with glmnet.cv glmnet.obj selectecVar reformat_dataSet
  res <- list(glmnet.cv=glmnet.cv, glmnet.obj=glmnet.obj, glmnet.cox=glmnet.cox, 
              nonzero.coef=nonzero.coef, selectedVar=selectedVar, 
              reformat_dataSet=reformat_dataSet
              # glmnet.cox.meaningful=glmnet.cox.meaningful, selectedVarCox=selectedVarCox
              )
  return(res)
}
 
 
## This function takes as input a surv_obj (testdate Surv object) and a testdata prediction matrix. Calculates a Brier score
## for either a cox model or a glmnet. Returns an ipec object which is the Brier score (requires peperr and c060)
 
brier_calc <- function(surv_obj, predict_matrix, model=c("cox","glmnet")){
  
  if(model == "cox"){
    peperr <- peperr.cox <- peperr(response=surv_obj, x=predict_matrix,
                                   fit.fun=fit.coxph, load.all=TRUE,
                                   indices=resample.indices(n=nrow(surv_obj), method="boot",
                                                            sample.n=50))      
  }else if(model == "glmnet"){
    peperr <- peperr(response=surv_obj, x=predict_matrix,
                     fit.fun=fit.glmnet, args.fit=list(family="cox"),
                     complexity=complexity.glmnet,args.complexity=list(family="cox"), load.all=TRUE,
                     indices=resample.indices(n=nrow(surv_obj), method="boot", sample.n=50))
  }
 
  # Get error predictions from peperr
  prederr <- perr(peperr)
 
  # Integrated prediction error Brier score calculation
  ipec <- ipec(prederr, eval.times=peperr$attribute, response=surv_obj)
 
  res <- list(peperr=peperr, ipec=ipec)
  return(res)
}
 
 
trainset <- dataset2predictmatrix(dataset)
models <- glmnetfun(trainset$predict_matrix, trainset$surv_obj)
# rfsrc <- rfsrcfunc(dataset)
briercox <- brier_calc(trainset$surv_obj, models$selectedVar, model="cox")
# requires c060 package which doesn't work on Linux
brierglmnet <- brier_calc(trainset$surv_obj, trainset$predict_matrix, model="glmnet")


# briermeaningful <- brier_calc(trainset$surv_obj, models$selectedVarCox, model="cox")
# brierRFSRC <- pec_calc(rfsrc$obj, dataset)
# brierRfit <- pec_calc(rfsrc$rfit, rfsrc$selectedDataset)
# # calculate brier score for coxph with feature selection based on rfsrc
# selectedDataset <- dataset2predictmatrix(rfsrc$selectedDataset)
# brierCox.rfsrc <- brier_calc(selectedDataset$surv_obj, selectedDataset$predict_matrix, model="cox")

# save image of the workspace after each iteration
save.image(file)
 
## C-Index calculation 50 iter bootstrapping
## initialization of lists for various outputs
cIndexCox <- list()
cIndexglmnet <- list ()
# cIndexCoxAIC <- list()
# cIndexRFSRC <- list()
# cIndexRfit <- list()
# cIndexCox.rfsrc <- list()
 
i <- 1
# pb <- txtProgressBar(min = 1, max = 50, style = 3) ## initialize a progress bar
train_results_list <- list()
# rfsrc_train_list <- list()
 
 
while (i < 51){
  
  print(paste("Iteration:",i))
# setTxtProgressBar(pb, i)## run progress bar
  train <- sample(1:nrow(dataset), nrow(dataset), replace = TRUE) ## random sampling with replacement
# create a dataframe for trainSet with time, status and selected variables in binary representation for evaluation in pec
  trainset <- dataset[train,]
  testset <- dataset[-train,]
  train_matrix <- dataset2predictmatrix(trainset)
 
  train_results <- glmnetfun(train_matrix$predict_matrix, train_matrix$surv_obj) #runs the sequence of glmnet.cv->glmnet.obj->glmnet.cox
  train_results_list <- c(train_results_list, list(train_results)) #create a list of results of all runs
 
  # compute c-index (Harrell's) for cox-glmnet models
  if (!is.null(train_results$glmnet.cox)){
    # creates a matrix of only the variables identified by the glmnet
    test_matrix <- dataset2predictmatrix(testset, nonzero.coef=train_results$nonzero.coef)
    reformat_testSet <- as.data.frame(cbind(test_matrix$surv_obj, test_matrix$predict_matrix)) #casts matrix to dataframe
    cIndexCox <- c(cIndexCox, 1-rcorr.cens(predict(train_results$glmnet.cox, 
                                                   reformat_testSet), test_matrix$surv_obj)[1])
    i=i+1
    # compute c-index (Harrell's) for glmnet models
    test_matrix <- dataset2predictmatrix (testset) # create matrix of the entire testset for glmnet evaluation
    cIndexglmnet <- c(cIndexglmnet, 1-rcorr.cens(predict(train_results$glmnet.cv,                                                      test_matrix$predict_matrix),test_matrix$surv_obj)[1])
 
    save.image(file)
  }
 
}

meanCICox <- c(mean(unlist(cIndexCox), na.rm=TRUE), std.error(unlist(cIndexCox)))
meanCIglmnet <- c(mean(unlist(cIndexglmnet), na.rm=TRUE), std.error(unlist(cIndexglmnet)))
```
