---
title: "Obtain the uncommon data for validation"
date: "Created: 2021-01-30; updated: `r Sys.Date()`"
author: 
  - name: "Hua Zou"
    email: "zouhua1@outlook.com"
output: 
  html_notebook:
    codes: hide
---

### Introduction

All the uncommon data will be used for validation of the model

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(data.table)
library(impute)

rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)
```


### load phenotype
```{r}
phen <- fread("../../Result/phenotype/all_survival_data.tsv")
com_phen <- fread("../../Result/phenotype/common_survival_data.tsv")

prof <- fread("../../Study/RawData/1080samples_expression_rawdata.txt")
```


### curation

#### get the uncommon samples information
```{r}
uncom_phen <- phen %>% filter(!Barcode%in%com_phen$Barcode)
```



#### filter expression matrix

* First, the biological features (e.g. genes/miRNAs) were removed if having zero value in more than 20% of patients. 

* The samples were removed if missing across more than 20% features. 

* Then we used impute function from R impute package, to fill out the missing values. 

* Lastly, we removed input features with zero values across all samples.

```{r}
get_profile <- function(profile    = prof,
                        metadata   = uncom_phen,
                        occurrence = 0.2,
                        present    = 0.2){
  # profile    = prof
  # metadata   = uncom_phen
  # occurrence = 0.2
  # present    = 0.2
  
  colnames(profile) <- c(colnames(profile)[1:3], gsub("-", "_", colnames(profile)[4:ncol(profile)]))
  
  sid <- intersect(uncom_phen$Barcode, colnames(profile))
  datprf <- profile %>% dplyr::select(c(colnames(profile)[1:3], sid))

  filter_fill_fun <- function(tag){
    
    # tag = data_type[1]
    
    dat <- datprf %>% filter(Platform == tag) %>%
                dplyr::select(-c("Platform", "Description"))
    # feature occurrence
    feature_occ <- apply(dat[, -1], 1, function(x){sum(x[!is.na(x)] != 0)/length(x)}) %>% 
      data.frame() %>% 
      setNames("feature_occurrence") %>%
      mutate(GeneSymbol = dat$GeneSymbol)
    remain_feature <- feature_occ %>% filter(feature_occurrence > occurrence)
    
    # sample occurrence 
    sample_occ <- apply(dat[, -1], 2, function(x){sum(x[!is.na(x)] != 0)/length(x)}) %>% 
      data.frame() %>% 
      setNames("sample_occurrence") %>%
      rownames_to_column("sample")
    remain_sample <- sample_occ %>% filter(sample_occurrence > present)
    
    # remove duplicate features
    dat.cln <- dat %>% filter(GeneSymbol%in%as.character(remain_feature$GeneSymbol)) %>%
      dplyr::select(c("GeneSymbol", as.character(remain_sample$sample))) 
    dat.cln$median <- apply(dat.cln[, -1], 1, median)
    dat.cln <- with(dat.cln, dat.cln[order(GeneSymbol, median, decreasing = T), ])
    dat.symbol <- dat.cln[!duplicated(dat.cln$GeneSymbol), ] %>%
      dplyr::select(-median) %>% 
      column_to_rownames("GeneSymbol")
    
    # impute.knn 
    set.seed(12345)
    saved.state <- .Random.seed
    dat.imputed <- impute.knn(as.matrix(dat.symbol))
    
    return(dat.imputed)    
  }
  
  data_type <- unique(datprf$Platform)
  #print(data_type)
  copyNumber <- filter_fill_fun(data_type[1])
  geneExp <- filter_fill_fun(data_type[2])
  methylation <- filter_fill_fun(data_type[3])
  protein_RPPA <- filter_fill_fun(data_type[4])
  
  copyNumber_prf <- copyNumber$data
  geneExp_prf <- geneExp$data
  methylation_prf <- methylation$data
  protein_RPPA_prf <- protein_RPPA$data
  
  metadata_cln <- metadata %>% filter(Barcode%in%sid)
  
  res <- list(meta = metadata_cln,
              CN = copyNumber_prf,
              GE = geneExp_prf,
              ML = methylation_prf,
              PR = protein_RPPA_prf)
  
  return(res)
}

data_trim_res <- get_profile(profile = prof,
                             metadata = uncom_phen)
```


### output
```{r}
dir_prf <- "../../Result/profile"
if(!dir.exists(dir_prf)){
  dir.create(dir_prf)
}

# copyNumber
copyNumber.path <- paste(dir_prf, "uncommon_copyNumber_filter.tsv", sep = "/")
write.table(data_trim_res$CN, file = copyNumber.path, quote = F, sep = "\t")

# geneExp
geneExp.path <- paste(dir_prf, "uncommon_geneExp_filter.tsv", sep = "/")
write.table(data_trim_res$GE, file = geneExp.path, quote = F, sep = "\t")

# methylation
methylation.path <- paste(dir_prf, "uncommon_methylation_filter.tsv", sep = "/")
write.table(data_trim_res$ML, file = methylation.path, quote = F, sep = "\t")

# protein_RPPA
protein_RPPA.path <- paste(dir_prf, "uncommon_protein_RPPA_filter.tsv", sep = "/")
write.table(data_trim_res$PR, file = protein_RPPA.path, quote = F, sep = "\t")


dir_phe <- "../../Result/phenotype"
if(!dir.exists(dir_phe)){
  dir.create(dir_phe)
}


# total 
uncommon_survival.path <- paste(dir_phe, "uncommon_survival_data.tsv", sep = "/")
write.table(data_trim_res$meta, file = uncommon_survival.path, quote = F, sep = "\t", row.names = F)
```


### version
```{r}
sessionInfo()
```


### Reference 

1. [TCGA数据下载整理问题](https://www.jianshu.com/p/a2dd6cb7ab13)

1. [fill out the missing values](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2432074/)
