---
title: "Curate data"
date: "Created: 2021-01-06; updated: `r Sys.Date()`"
author: 
  - name: "Hua Zou"
    email: "zouhua1@outlook.com"
output: 
  html_notebook:
    codes: hide
---

### Introduction

We first used TCGA-Assembler Version 2 with the keywords **KIRC** to search and process the four-omics data (CNVs; mRNA; DNA methylation; RPPA), and then performed the next precedure to filter the unmatched data. 

The original script for downloading the RawData was here *KIRC/Study/Script/data-KIRC.R*

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(dplyr)
library(tibble)
library(data.table)
library(ggplot2)
library(impute)

rm(list = ls())
options(stringsAsFactors = F)
options(future.globals.maxSize = 1000 * 1024^2)
```



### load phenotype
```{r}
phen <- read.csv("../../Study/Phenotype/KIRC_clinical__information.csv")
prof <- fread("../../Study/RawData/290samples_expression.txt")
```


### curation

#### choose index of the phenotype for survival analysis
```{r}
phen.cln <- phen %>% 
  mutate(stage_event_tnm_categories = paste0(pathologic_T, pathologic_N, pathologic_M)) %>%
  dplyr::select(bcr_patient_barcode, gender, vital_status, days_to_death, 
                          days_to_last_followup, race, person_neoplasm_cancer_status, 
                          age_at_initial_pathologic_diagnosis, laterality, 
                          neoplasm_histologic_grade, pathologic_stage, stage_event_tnm_categories) %>%
  distinct(bcr_patient_barcode, .keep_all = TRUE)

# dead_patient 
dead_patient <- phen.cln %>%
  dplyr::filter(vital_status == "Dead") %>%
  dplyr::select(-days_to_last_followup) %>%
  reshape::rename(c(bcr_patient_barcode = "Barcode",
           gender = "Gender",
           vital_status = "OS",
           days_to_death = "OS.Time",
           race = "Race",
           person_neoplasm_cancer_status = "cancer_status",
           age_at_initial_pathologic_diagnosis = "Age",
           neoplasm_histologic_grade = "Grade",
           pathologic_stage = "Stage",
           stage_event_tnm_categories = "TNM")) %>%
  mutate(OS = ifelse(OS == "Dead", 1, 0)) %>%
  mutate(OS.Time = as.numeric(as.character(OS.Time))/365)

# alive_patient 
alive_patient <- phen.cln %>%
  dplyr::filter(vital_status == "Alive") %>%
  dplyr::select(-days_to_death) %>%
  reshape::rename(c(bcr_patient_barcode = "Barcode",
           gender = "Gender",
           vital_status = "OS",
           days_to_last_followup = "OS.Time",
           race = "Race",
           person_neoplasm_cancer_status = "cancer_status",
           age_at_initial_pathologic_diagnosis = "Age",
           neoplasm_histologic_grade = "Grade",
           pathologic_stage = "Stage",
           stage_event_tnm_categories = "TNM")) %>%
  mutate(OS = ifelse(OS == "Dead", 1, 0)) %>%
  mutate(OS.Time = as.numeric(as.character(OS.Time))/365)


# combine data 
survival_data <- rbind(dead_patient, alive_patient) %>%
  mutate(Barcode = gsub("-", "_", Barcode))
```



#### filter expression matrix

* First, the biological features (e.g. genes/miRNAs) were removed if having zero value in more than 20% of patients. 

* The samples were removed if missing across more than 20% features. 

* Then we used impute function from R impute package, to fill out the missing values. 

* Lastly, we removed input features with zero values across all samples.

```{r}
get_profile <- function(profile    = prof,
                        metadata   = survival_data,
                        occurrence = 0.2,
                        present    = 0.2){
  # profile    = prof
  # metadata   = survival_data
  # occurrence = 0.2
  # present    = 0.2
  
  colnames(profile) <- c(colnames(profile)[1:3], gsub("-", "_", colnames(profile)[4:ncol(profile)]))

  filter_fill_fun <- function(tag){
    
    # tag = data_type[1]
    
    dat <- profile %>% filter(Platform == tag) %>%
                dplyr::select(-c("Platform", "Description"))
    # feature occurrence
    feature_occ <- apply(dat[, -1], 1, function(x){sum(x[!is.na(x)] != 0)/length(x)}) %>% 
      data.frame() %>% 
      setNames("feature_occurrence") %>%
      mutate(GeneSymbol = dat$GeneSymbol)
    remain_feature <- feature_occ %>% filter(feature_occurrence > occurrence)
    
    # sample occurrence 
    sample_occ <- apply(dat[, -1], 2, function(x){sum(x[!is.na(x)] != 0)/length(x)}) %>% 
      data.frame() %>% 
      setNames("sample_occurrence") %>%
      rownames_to_column("sample")
    remain_sample <- sample_occ %>% filter(sample_occurrence > present)
    
    # remove duplicate features
    dat.cln <- dat %>% filter(GeneSymbol%in%as.character(remain_feature$GeneSymbol)) %>%
      dplyr::select(c("GeneSymbol", as.character(remain_sample$sample))) 
    dat.cln$median <- apply(dat.cln[, -1], 1, median)
    dat.cln <- with(dat.cln, dat.cln[order(GeneSymbol, median, decreasing = T), ])
    dat.symbol <- dat.cln[!duplicated(dat.cln$GeneSymbol), ] %>%
      dplyr::select(-median) %>% 
      column_to_rownames("GeneSymbol")
    
    # impute.knn 
    set.seed(12345)
    saved.state <- .Random.seed
    dat.imputed <- impute.knn(as.matrix(dat.symbol))
    
    return(dat.imputed)    
  }
  
  data_type <- unique(profile$Platform)
  #print(data_type)
  copyNumber <- filter_fill_fun(data_type[1])
  geneExp <- filter_fill_fun(data_type[2])
  methylation <- filter_fill_fun(data_type[3])
  protein_RPPA <- filter_fill_fun(data_type[4])
  
  copyNumber_prf <- copyNumber$data
  geneExp_prf <- geneExp$data
  methylation_prf <- methylation$data
  protein_RPPA_prf <- protein_RPPA$data
  
  common_sampleid <- intersect(metadata$Barcode, 
                        intersect(colnames(copyNumber_prf), 
                          intersect(colnames(geneExp_prf), 
                            intersect(colnames(methylation_prf), colnames(protein_RPPA_prf))
                          )
                        )
                                  
                      )
  
  metadata.cln <- metadata %>% filter(Barcode %in% common_sampleid)
  copyNumber_prf.cln <- copyNumber_prf[, colnames(copyNumber_prf)%in%common_sampleid]
  geneExp_prf.cln <- geneExp_prf[, colnames(geneExp_prf)%in%common_sampleid]
  methylation_prf.cln <- methylation_prf[, colnames(methylation_prf)%in%common_sampleid]
  protein_RPPA_prf.cln <- protein_RPPA_prf[, colnames(protein_RPPA_prf)%in%common_sampleid]
  
  
  res <- list(meta = metadata.cln, 
              CN = copyNumber_prf.cln,
              GE = geneExp_prf.cln,
              ML = methylation_prf.cln,
              PR = protein_RPPA_prf.cln)
  
  return(res)
}

data_trim_res <- get_profile(profile = prof,
                             metadata = survival_data)
```


### output
```{r}
dir_prf <- "../../Result/profile"
if(!dir.exists(dir_prf)){
  dir.create(dir_prf)
}

# copyNumber
copyNumber.path <- paste(dir_prf, "copyNumber_filter.tsv", sep = "/")
write.table(data_trim_res$CN, file = copyNumber.path, quote = F, sep = "\t")

# geneExp
geneExp.path <- paste(dir_prf, "geneExp_filter.tsv", sep = "/")
write.table(data_trim_res$GE, file = geneExp.path, quote = F, sep = "\t")

# methylation
methylation.path <- paste(dir_prf, "methylation_filter.tsv", sep = "/")
write.table(data_trim_res$ML, file = methylation.path, quote = F, sep = "\t")

# protein_RPPA
protein_RPPA.path <- paste(dir_prf, "protein_RPPA_filter.tsv", sep = "/")
write.table(data_trim_res$PR, file = protein_RPPA.path, quote = F, sep = "\t")


dir_phe <- "../../Result/phenotype"
if(!dir.exists(dir_phe)){
  dir.create(dir_phe)
}

# total 
all_survival.path <- paste(dir_phe, "all_survival_data.tsv", sep = "/")
write.table(survival_data, file = all_survival.path, quote = F, sep = "\t", row.names = F)

# total 
common_survival.path <- paste(dir_phe, "common_survival_data.tsv", sep = "/")
write.table(data_trim_res$meta, file = common_survival.path, quote = F, sep = "\t", row.names = F)
```


### version
```{r}
sessionInfo()
```


### Reference 

1. [TCGA数据下载整理问题](https://www.jianshu.com/p/a2dd6cb7ab13)

1. [fill out the missing values](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2432074/)
